Exploring convergence issues beyond 4-CPL for SPD of shoreline dates

### 2023-08-15 analysis/paper/paper.Rmd
Setting up supp 2

# Introduction
Variation in population patterns are regarded as important drivers and potential consequences of cultural variation, and is thus of critical importance to our understanding of past human societies [e.g. @french2021; @shennan2000; @shennan2021]. During the last decade, the frequency distribution of radiocarbon dates have become the preferred proxy for archaeologists studying prehistoric population dynamics [@crema2022; @french2021; @shennan2013], owing in part to the fact that radiocarbon dates represent the most temporally and spatially widespread archaeological data available for studying past population densities [@bird2022]. This has also been the case for Norwegian Stone Age archaeology [@solheim2018; @solheim2020a; @nielsen2019; @nielsen2021; @bergsvik2021; @jorgensen2020; @lundstrom2023; see also @persson2009].

@@

To what degree the radiocarbon record is determined by past population numbers might vary geographically and chronologically, based on variation in investigatory and taphonomic factors [@surovell2009b; @bluhm2019], as well as cultural processes within prehistoric populations. One example of the latter is the difference that might exist between farmer and forager populations, where @freeman2018 have suggested that an increased per capita energy consumption introduced with farming means that ^14^C-dates should not be weighted equally when making relative population estimates across such populations. Similarly, while site counts have also been invoked for the analysis of past population dynamics, these are likely to be impacted by factors such as land-use and mobility patterns, and settlement nucleation and dispersion, as well as variation in archaeological investigations and what has been taken to constitute a 'site' [@palmisano2017]. However, these could be considered theoretical issues that have implications for how fluctuations in these proxies should be interpreted, and possibly weighted differently according some criteria. Before any such fluctuations are given any substantive interpretation, however, there is a host of methodological issues that have to be considered [see @carleton2021; @crema2022 for reviews].

The most critical of these follow from the fact that the summation of the probabilities associated with the dates for SPDs is not a directly coherent statistical procedure. This is because the summed probabilities can no longer been seen as probabilities, but rather represent the combination of events and chronological uncertainties, making the two indistinguishable, and rendering the interpretation of the resulting sum difficult [@blackwell2003; @crema2022]. As @timpson2021[2] put it: 'the SPD is not *the* single best explanation of the data, nor even *a* single explanation of the data, but rather a conflation of many possible explanations simultaneously, each of which is mired by the artefacts inherited from the calibration wiggles.' The SPD is not a model. It is the combined representation of a range of possible explanations for the data -- the frequency of dated events combined with the variable uncertainty associated with these [@carleton2021]. This means that a direct visual inspection of SPDs can be misleading, that they cannot be directly analysed to draw inferences on population dynamics, and that they cannot be directly compared to other time-series data [@carleton2018; @timpson2021; @crema2022]. While this problem can never be entirely resolved, a range of approaches have been developed in an attempt to work around this issue.

@@

The procedure outlined above represent the standard approach in efforts to model population dynamics based on RSPDs. This can be helpful for getting an impression of the overall developments of population proxies, and is an especially useful first step in the treatment of shoreline dated sites, as few assumptions concerning this overall development could be made. However, standard models frequently used to evaluate RSPDs are consistently being rejected in the literature and, it is worth underscoring, that deviations from these rejected null models do not in themselves give statistical justification of the interpretation of any interpretations of the local deviations, nor a subsequent direct analysis of the SPDs themselves -- statistical or otherwise.

@@

The borders of the municipalities within which the shoreline dated sites are located were used to limit the radiocarbon sample. Radiocarbon dates were taken from @roalkvam2023 and @solheim. Dates done on food crusts were then excluded due the issue of marine reservoir effects [@nielsen2019, 83]. Following both from the point made by @freeman2018 concerning the comparison of radiocarbon dates from populations with different economic modes, as noted above, and from the fact that shoreline dating appears less reliable after 4000 BCE and appears to loose its utility c. 2500 BCE, the Mesolithic is the main focus of this analysis. Consequently, probabilities falling later than 2500 BCE were excluded from the analysis, leading to a final sample of 310 ^14^C-dates. Probabilities falling after 4000 BCE, and the first possible small-scale introduction of agriculture in south-eastern Norway, were retained to account for edge-effects but these results should be treated with care for the reasons noted above.

@@

With an estimated trajectory of shoreline displacement at the site to be dated, the elevation of the site can be drawn on to find a likely date for when it was in use, under the assumption that it was in use when located on or close to the shoreline. Figure \@ref(fig:shoredate)C indicates where the mean elevation of the polygon representing the example site intersects the interpolated displacement curve. As it can reasonably be assumed that a site was not in use when located under water, this effectively represents a *terminus post quem* (TPQ) date. One approach to shoreline dating has been to find where the elevation of the site intersects the mean of the upper and lower limit of the displacement curve [e.g. @breivik2014]. However, TPQ dates that do not account for the likely distance between site and sea when the site was in use limits the further inferential steps that can be taken. Furthermore, the common practice of adding a constant error to a point-estimate, as mentioned above, does not take into account that the rate of shoreline displacement varies both spatially and temporally, and therefore introduces bias to the result on both the spatial and temporal scale, while also overestimating the precision of the method in most cases.

@@

```{r shoredate, echo = FALSE, fig.cap = "A) An example site (Askeladden ID 111967-1) relative to the isobases of the displacement curves.  B) The geologically derived shoreline displacement curves for the region and the curve interpolated to the example site. C) The black line on the y-axis represents the elevation of the site. This crosses the interpolated displacement curve at 7680 and 7490 BCE, which thus in practice gives an estimate of when the location of the site emerged from the sea, giving a \\textit{terminus post quem} date. D) The shoreline dating approach used for this study, where the gamma distrbiution on the y-axis describes the likely vertical relationship between the shoreline and the site when it was in use.",  out.width = "500px", fig.align="center", regfloat = TRUE}

@@

When the frequency distribution of shoreline dates have been investigated in the past, this has typically involved finding point-estimates of dates by either using the mean of displacement curves, or by taking mean of the TPQ dates for each site (as illustrated in Figure \@ref(fig:shoredate)C). As outlined above,  these point-estimates are then typically aggregated in bins of 200 or 500 years. This approach is illustrated in Figure \@ref(fig:sumshore)B. The SSPD resulting from using the probabilistic method of shoreline dating, given in Figure \@ref(fig:sumshore)C, demonstrates that accounting for the site-sea relationship gives a substantially different result compared to the more traditional approach, which is clearly indicated by the shifts in the positions of peaks and troughs in the distribution after c. 8000 BCE.

@@

Although there are generally few ^14^C-dates from the earliest part of the Mesolithic in Norway [e.g. @åstveit2018; @kleppe2018; @breivik2018a; @solheim2018], it is established that Stone Age sites in southern Norway date to the period from before the earliest ^14^C-dates in the study area for this paper, and at least from as far back as around 9300 BCE [e.g. @damlien2018b; @glorstad2016]. While the SSPD achieved here starts from c. 9500 BCE, the RSPD shows no evidence of human activity until c. 8500 BCE. What has caused the lack of ^14^C-dates from this earliest period has been given many possible explanations through the years. These range from taphonomic loss to diverse environmental and/or cultural factors, such as the use of shrubs or rapidly deteriorating seal blubber as fuel for burning, or a preference for consuming raw, dried or fermented food [see @bjerck2017; @damm2022]. A possible lack of reliable access to firewood has also been seen as possible contributor to this pattern, although evidence for the early presence of forested landscapes, at least in the southern parts of the country, has now rendered this a less favoured hypothesis [@solheim2018; cf. @damm2022; @sørensen2014]. While taphonomic and ecological factors are likely part of the picture, we view it as probable that the rapid growth at the start of RSPD and the magnitude of the SSPD in the period preceding this suggest that cultural and demographic factors are central drivers behind this pattern.

@@

Following this earliest phase of rapid increase in shoreline dated sites, there is a drop from this initial peak, starting from around 8800 BCE, concluding around 8500 BCE and remaining within the simulation envelope until c. 7900 BCE. It has previously been suggested that a change in mobility patterns occurs some time around 8500--8300 BCE, involving a reduction in residential mobility [e.g. @damlien2018b; @roalkvam2022]. Furthermore, both genetic data and technological analyses of lithic inventories indicate that the initial phase of human colonisation from the south is followed by an influx, and possible mixing of people from the east around the same time [@günther2018; @manninen2021]. If this narrative is coupled with the data observed here, these migratory events could appear to result in a relatively sudden drop in site-count as indicated in the SSPD, which in turn could be related to a change in mobility patterns. The end of this drop also roughly corresponds to the first ^14^C-dates in the RSPD around 8500 BCE, leading to the logistic growth that plateaus in the first centuries after 8000 BCE. The plateau of the logistic growth in the RSPD also has a corresponding positive deviation in the SSPD from the exponential model -- an increase that was also noted by @berg-hansen2022 -- which could indicate that this demographic signal is reflected in the frequency of shoreline dated sites as well.

@@

Directly relating these patterns to demographic developments or land-use and mobility patterns is challenging given the data available here. Given the stationarity of the logistic model of the RSPD after c. 7500 BCE, it is possible that fluctuations in the SSPD reflect variation in mobility patterns that are unrelated to demographic developments. There appears to be a negative deviation from the logistic null model in the RSPD corresponding to the one in the SSPD just before 6000 BCE. However, the logistic model could not be rejected for the RSPD, and so it is inappropriate to interpret any deviations from the Monte Carlo envelope and place any analytical weight on the deviations in the RSPD, at least given the present data. While it is possible that the fairly limited sample of ^14^C-dates has led to an underestimation of this deviation, there is, additionally, no indications in the RSPD of an increased demographic signal in the period just before 5000 BCE [see also @solheim2018].

>>>>

# Introduction
Variation in population patterns are regarded as important drivers and potential outcomes of cultural variation, and is thus of critical importance to our understanding of past human societies [e.g. @french2021; @shennan2000; @shennan2021]. During the last decade, the frequency distribution of radiocarbon dates have become the preferred proxy for archaeologists studying prehistoric population dynamics [@crema2022; @french2021; @shennan2013], owing in part to the fact that radiocarbon dates represent the most temporally and spatially widespread archaeological data available for studying past population densities [@bird2022]. This has also been the case for Norwegian Stone Age archaeology [@solheim2018; @solheim2020a; @nielsen2019; @nielsen2021; @bergsvik2021; @jorgensen2020; @lundstrom2023; see also @persson2009].

@@

To what degree the radiocarbon record is determined by past population numbers might vary geographically and chronologically, based on variation in investigatory and taphonomic factors [@surovell2009b; @bluhm2019], as well as cultural processes within prehistoric populations. One example of the latter is the difference that might exist between farmer and forager populations, where @freeman2018 have suggested that an increased per capita energy consumption introduced with farming means that ^14^C-dates should not be weighted equally when making relative population estimates across such populations. Similarly, while site counts have also been invoked for the analysis of past population dynamics, these are likely to be impacted by factors such as land-use and mobility patterns, and settlement nucleation and dispersion, as well as variation in archaeological investigations and what has been taken to constitute a site [@drennan2015, 16--20; @palmisano2017]. However, these could be considered theoretical issues that have implications for how fluctuations in these proxies should be interpreted, and possibly weighted differently according some criteria. Before any such fluctuations are given any substantive interpretation, however, there is a host of methodological issues that have to be considered [see @carleton2021; @crema2022 for reviews].

The most critical of these follow from the fact that the summation of the probabilities associated with the dates for SPDs is not a directly coherent statistical procedure. This is because the summed probabilities can no longer been seen as probabilities, but rather represent the combination of events and chronological uncertainties, making the two indistinguishable, and rendering the interpretation of the resulting sum difficult [@blackwell2003; @crema2022]. As @timpson2021[2] put it: 'the SPD is not *the* single best explanation of the data, nor even *a* single explanation of the data, but rather a conflation of many possible explanations simultaneously, each of which is mired by the artefacts inherited from the calibration wiggles.' The SPD is not a model. It is the combined representation of a range of possible explanations for the data -- the frequency of dated events combined with the variable uncertainty associated with these [@carleton2021]. This means that a direct visual inspection of SPDs can be misleading, that they cannot be directly analysed to draw inferences on population dynamics, and that they cannot be directly compared to other time-series data [@carleton2018; @timpson2021; @crema2022]. While this problem cannot be entirely resolved, a range of approaches have been developed in an attempt to work around this issue.

@@

The procedure outlined above represent the standard approach in efforts to model population dynamics based on RSPDs. This can be helpful for getting an impression of the overall developments of population proxies, and is an especially useful first step in the treatment of shoreline dated sites, as few assumptions concerning this overall development could be made. However, standard models frequently used to evaluate RSPDs are consistently being rejected in the literature and, it is worth underscoring, deviations from these rejected null models do not in themselves give statistical justification for the interpretation of any local deviations, nor a subsequent direct analysis of the SPDs themselves.

@@

The borders of the municipalities within which the shoreline dated sites are located were used to limit the radiocarbon sample. Radiocarbon dates were taken from @roalkvam2023 and an unpublished data set collated by Solheim and Kjetil Loftsgarden. Dates done on food crusts were then excluded due the issue of marine reservoir effects [@nielsen2019, 83]. Following both from the point made by @freeman2018 concerning the comparison of radiocarbon dates from populations with different economic modes, as noted above, and from the fact that shoreline dating appears less reliable after 4000 BCE and appears to loose its utility c. 2500 BCE, the Mesolithic is the main focus of this analysis. Consequently, probabilities falling later than 2500 BCE were excluded from the analysis, leading to a final sample of 310 ^14^C-dates. Probabilities falling after 4000 BCE, and the first possible small-scale introduction of agriculture in south-eastern Norway, were retained to account for edge-effects but these results should be treated with care for the reasons noted above.

@@

With an estimated trajectory of shoreline displacement at the site to be dated, the elevation of the site can be drawn on to find a likely date for when it was in use, under the assumption that it was in use when located on or close to the shoreline. Figure \@ref(fig:shoredate)C indicates where the mean elevation of the polygon representing the example site intersects the interpolated displacement curve. As it can reasonably be assumed that a site was not in use when located under water, this effectively represents a *terminus post quem* (TPQ) date. However, TPQ dates that do not account for the likely distance between site and sea when the site was in use limits the further inferential steps that can be taken. Furthermore, the common practice of adding a constant error or binning point-estimates of shoreline dates, as mentioned above, does not take into account that the rate of shoreline displacement varies both spatially and temporally, and therefore introduces bias to the result on both the spatial and temporal scale, while also overestimating the precision of the method in most cases.

@@

```{r shoredate, echo = FALSE, fig.cap = "A) An example site (Askeladden ID 111967-1) relative to the isobases of the displacement curves.  B) The geologically derived shoreline displacement curves for the region and the curve interpolated to the example site. C) The black line on the y-axis represents the elevation of the site. This crosses the interpolated displacement curve at 7680 and 7490 BCE, which thus in practice gives an estimate of when the location of the site emerged from the sea and represents a \\textit{terminus post quem} date for the site. D) The shoreline dating approach used for this study, where the gamma distrbiution on the y-axis describes the likely vertical relationship between the shoreline and the site when it was in use.",  out.width = "500px", fig.align="center", regfloat = TRUE}

@@

When the frequency distribution of shoreline dates have been investigated in the past, this has typically involved finding point-estimates of dates by either using the mean of displacement curves, or by taking mean of the TPQ dates for each site (as illustrated in Figure \@ref(fig:shoredate)C). As outlined above,  these point-estimates are then typically aggregated in bins of 200 or 500 years. This approach is illustrated in Figure \@ref(fig:sumshore)B. The SSPD resulting from using the probabilistic method of shoreline dating, given in Figure \@ref(fig:sumshore)C, demonstrates that accounting for the site-sea relationship gives a substantially different result compared to the more traditional approach, which is indicated by the shifts in the positions of peaks and troughs in the distribution after c. 8000 BCE.

@@

Although there are generally few ^14^C-dates from the earliest part of the Mesolithic in Norway [e.g. @åstveit2018; @kleppe2018; @breivik2018a; @solheim2018], it is established that Stone Age sites in southern Norway date to the period from before the earliest ^14^C-dates in the study area for this paper, and at least from as far back as around 9300 BCE [e.g. @damlien2018b; @glorstad2016]. While the SSPD achieved here starts from c. 9500 BCE, the RSPD shows no evidence of human activity until c. 8500 BCE. What has caused the lack of ^14^C-dates in south-eastern Norway from this earliest period has been given many possible explanations through the years. These range from taphonomic loss to diverse environmental and/or cultural factors, such as the use of shrubs or rapidly deteriorating seal blubber as fuel for burning, or a preference for consuming raw, dried or fermented food [see @bjerck2017; @damm2022]. A possible lack of reliable access to firewood has also been seen as possible contributor to this pattern, although evidence for the early presence of forested landscapes, at least in the southern parts of the country, has now rendered this a less favoured hypothesis [@solheim2018; cf. @damm2022; @sørensen2014]. While taphonomic and ecological factors are likely part of the picture, we view it as probable that the rapid growth at the start of RSPD and the magnitude of the SSPD in the period preceding this suggest that cultural and demographic factors are central drivers behind this pattern.

@@

Following this earliest phase of rapid increase in shoreline dated sites, there is a drop from this initial peak, starting from around 8800 BCE, concluding around 8500 BCE and remaining within the simulation envelope until c. 7900 BCE. It has previously been suggested that a change in mobility patterns occurs some time around 8500--8000 BCE, involving a reduction in residential mobility [e.g. @damlien2018b; @roalkvam2022]. Furthermore, both genetic data and technological analyses of lithic inventories indicate that the initial phase of human colonisation from the south is followed by an influx, and possible mixing of people from the east around the same time [@günther2018; @manninen2021]. If this narrative is coupled with the data observed here, these migratory events could appear to result in a relatively sudden drop in site-count as indicated in the SSPD, which in turn could be related to a change in mobility patterns. The end of this drop also roughly corresponds to the first ^14^C-dates in the RSPD around 8500 BCE, leading to the logistic growth that plateaus in the first centuries after 8000 BCE. The plateau of the logistic growth in the RSPD also has a corresponding positive deviation in the SSPD from the exponential model -- an increase that was also noted by @berg-hansen2022 -- which could indicate that this demographic signal is reflected in the frequency of shoreline dated sites as well.

@@

Directly relating these patterns to demographic developments or land-use and mobility patterns is challenging given the data available here. Given the stationarity of the logistic model of the RSPD after c. 7500 BCE, it is possible that fluctuations in the SSPD reflect variation in mobility patterns that are unrelated to demographic developments. There appears to be a negative deviation from the logistic null model in the RSPD corresponding to the one in the SSPD just before 6000 BCE. However, the logistic model could not be rejected for the RSPD, and so it is inappropriate to interpret any deviations from the Monte Carlo envelope and place any analytical weight on the deviations in the RSPD, at least given the present data. While it is possible that the fairly limited sample of ^14^C-dates has led to an underestimation of this deviation, there is, additionally, no indications in the RSPD of an increased demographic signal in the period just before 5000 BCE [see also @solheim2018]. A related point that is also worth making here is that the overall demographic trends that the modelling efforts can hope to identify are likely to have been subject to smaller scale demographic fluctuations that simply cannot be ascertained given the available data and its resolution [cf. @timpson2021; @tallavaara2021], and which impacts the comparison of the two proxies following both from the resolution of the dating methods and from the different sample sizes underlying the SPDs.

### 2023-08-15 analysis/supplementary/supplementary2.Rmd
Setting up supp 2

```{r data}
# Load data holding the PD array of shoreline dates
load(file = here::here("analysis/data/derived_data/shore_models_pd.RData"))

@@

```{r precompiled, echo = FALSE}
# For the compilation of this Rmarkdown document, the results were pre-compiled 
# and are loaded here. Set eval = TRUE in code chunks below to re-run the code.
load(file = here::here("analysis/data/derived_data/sampsize.RData"))

@@

  results[[i]] <- rbind(cpl4, cpl5, cpl6)
  save(results, temppds, sampsize,
       file = here::here("analysis/data/derived_data/sampsize.RData"))
}
```

@@

>>>>

```{r data}
# Load data holding the PD array of shoreline dates
load(file = here::here("analysis/data/derived_data/shore_pd.RData"))

@@

```{r precompiled, echo = FALSE}
# For the compilation of this Rmarkdown document, the results of the code chunk 
# below  were pre-compiled and are loaded here. Set eval = TRUE below to re-run
# the code.
load(file = here::here("analysis/data/derived_data/sampsize.RData"))

@@

  results[[i]] <- rbind(cpl4, cpl5, cpl6)
}
```

@@

# Different parameter settings for 5-CPL

The figure below displays the results of attempting to adjust parameter settings for improving the parameter search for the 5-CPL model. The identified model parameters are identical across the runs, the variation in the returned values is minute, and none of the searches reached convergence. The parameters used with `JDEoptim()` were altered one at a time, resulting in 11 separate searches. The following code was used as the starting point from where parameters were adjusted:
```{r, eval = FALSE}
# Base code from where the parameters were adjusted individually
JDEoptim(lower = rep(0, 9),
                   upper = rep(1, 9),
                   fn = objectiveFunction,
                   PDarray = pd,
                   type = 'CPL',
                   NP = 180)
```

```{r, echo = FALSE}

load(file = here::here("analysis/data/derived_data/exploring_cpl5.RData"))

cpl5iter <- convertPars(pars = cpl_5$par, years = minage:maxage, type = 'CPL')
cpl5iter$parameter <- "maxiter = 3600"
cpl5iter$value <- cpl_5$value
cpl5iter$convergence <- cpl_5$convergence

cpl5iter2 <- convertPars(pars = cpl_5iter2$par, years = minage:maxage, type = 'CPL')
cpl5iter2$parameter <- "maxiter = 7200"
cpl5iter2$value <- cpl_5iter2$value
cpl5iter2$convergence <- cpl_5iter2$convergence

cpl5np <- convertPars(pars = cpl_5np$par, years = minage:maxage, type = 'CPL')
cpl5np$parameter <- "NP = 360"
cpl5np$value <- cpl_5np$value
cpl5np$convergence <- cpl_5np$convergence

cpl5np2 <- convertPars(pars = cpl_5np2$par, years = minage:maxage, type = 'CPL')
cpl5np2$parameter <- "NP = 720"
cpl5np2$value <- cpl_5np2$value
cpl5np2$convergence <- cpl_5np2$convergence

cpl5jitter <- convertPars(pars = cpl_5jitter$par, years = minage:maxage, type = 'CPL')
cpl5jitter$parameter <- "jitter_factor = NULL"
cpl5jitter$value <- cpl_5jitter$value
cpl5jitter$convergence <- cpl_5jitter$convergence

cpl5jitter2 <- convertPars(pars = cpl_5jitter2$par, years = minage:maxage, type = 'CPL')
cpl5jitter2$parameter <- "jitter_factor = 0.01"
cpl5jitter2$value <- cpl_5jitter2$value
cpl5jitter2$convergence <- cpl_5jitter2$convergence

cpl5jitter3 <- convertPars(pars = cpl_5jitter3$par, years = minage:maxage, type = 'CPL')
cpl5jitter3$parameter <- "jitter_factor = 0.1"
cpl5jitter3$value <- cpl_5jitter3$value
cpl5jitter3$convergence <- cpl_5jitter3$convergence

cpl5jitter4 <- convertPars(pars = cpl_5jitter4$par, years = minage:maxage, type = 'CPL')
cpl5jitter4$parameter <- "jitter_factor = 0.0001"
cpl5jitter4$value <- cpl_5jitter4$value
cpl5jitter4$convergence <- cpl_5jitter4$convergence

cpl5mutprob1 <- convertPars(pars = cpl_5mutprob1$par, years = minage:maxage, type = 'CPL')
cpl5mutprob1$parameter <- "tau_pF = 0.5"
cpl5mutprob1$value <- cpl_5mutprob1$value
cpl5mutprob1$convergence <- cpl_5mutprob1$convergence

cpl5mutprob2 <- convertPars(pars = cpl_5mutprob2$par, years = minage:maxage, type = 'CPL')
cpl5mutprob2$parameter <- "tau_pF = 0.2"
cpl5mutprob2$value <- cpl_5mutprob2$value
cpl5mutprob2$convergence <- cpl_5mutprob2$convergence

cpl5mutprob3 <- convertPars(pars = cpl_5mutprob3$par, years = minage:maxage, type = 'CPL')
cpl5mutprob3$parameter <- "tau_pF = 1"
cpl5mutprob3$value <- cpl_5mutprob3$value
cpl5mutprob3$convergence <- cpl_5mutprob3$convergence

cpl5models <- rbind(cpl5iter, cpl5iter2, cpl5np, cpl5np2, cpl5jitter,
                    cpl5jitter2, cpl5jitter3, cpl5jitter4, cpl5mutprob1,
                    cpl5mutprob2, cpl5mutprob3)

plt1 <- ggplot() + geom_line(aes(x = as.numeric(rownames(SPD)), y = SPD[,1])) + geom_line(data = cpl5models, aes(x = year, y = pdf, col = parameter)) +
  labs(title = "Different parameter settings for 5-CPL search") + scale_x_reverse() +
  theme_bw()
plt2 <- ggplot() + geom_point(data = cpl5models, aes(x = value, y = parameter)) + geom_vline(xintercept = cpl_5$value, col = "red") +
  ylab("") +
  theme_bw()
plt3 <- ggplot() + geom_point(data = cpl5models, aes(x = as.factor(convergence), y = parameter, col = as.factor(convergence))) +
  ylab("") +
  theme_bw() + theme(legend.position =  "none", axis.text.y=element_blank(), axis.ticks.y=element_blank())

plt1

```{r, echo = FALSE}
plt2 + plt3
```
